Heuristics and Really Hard Instances for Subgraph Isomorphism Problems

### 2016-01-18 paper/paper.tex
Tweaks

For subgraph isomorphism we identify relevant control parameters: we can independently alter the
edge probability of the pattern graph, the edge probability of the target graph, and the relative
orders (number of vertices) of the pattern and target graphs.  For non-induced isomorphisms, with
the correct choice of parameters we see results very similar to those observed with boolean
satisfiability problems: there is a phase transition from satisfiable to unsatisfiable (and we can
predict its location), and we see a complexity peak occur with three different solvers near this
phase transition.

For certain choices of parameters for induced isomorphisms, there are two phase transitions, going
from satisfiable to unsatisfiable, and then from unsatisfiable back to satisfiable. Again, when
going from satisfiable to unsatisfiable (from either direction), instances go from being trivial to
really hard to solve. However, each of the three solvers we tested also finds the central
unsatisfiable region to be really hard---this may be unexpected, since this region is not near a
phase transition. To show that this is not a simple weakness of current subgraph isomorphism
algorithms, we verify that this region is also hard when using use a pseudo-boolean encoding, and
under reduction to the clique problem. Interestingly, the constrainedness measure
\citep{Gent:1996:Kappa} \emph{does} predict this difficult region---these instances provide (?? the
first??) evidence in favour of constrainedness, rather than proximity to a phase transition, being
an accurate predictor of difficulty.

>>>>

For subgraph isomorphism we identify three relevant control parameters: we can independently alter
the edge probability of the pattern graph, the edge probability of the target graph, and the
relative orders (number of vertices) of the pattern and target graphs.  For non-induced
isomorphisms, with the correct choice of parameters we see results very similar to those observed
with boolean satisfiability problems: there is a phase transition from satisfiable to unsatisfiable
(and we can predict its location), and we see a complexity peak occur with three different solvers
near this phase transition.

For certain choices of parameters for induced isomorphisms, there are two phase transitions, going
from satisfiable to unsatisfiable, and then from unsatisfiable back to satisfiable. Again, when
going from satisfiable to unsatisfiable (from either direction), instances go from being trivial to
really hard to solve. However, each of the three solvers we tested also finds the central
unsatisfiable region to be hard---this may be unexpected, since this region is not near a
phase transition. To show that this is not a simple weakness of current subgraph isomorphism
algorithms, we verify that this region is also hard when using use a pseudo-boolean encoding, and
under reduction to the clique problem. Interestingly, the constrainedness measure
\citep{Gent:1996:Kappa} \emph{does} predict this difficult region---these instances provide (?? the
first??) evidence in favour of constrainedness, rather than proximity to a phase transition, being
an accurate predictor of difficulty.
